### 2019-01-14 Paper/Bibliography.tex
References and other edits
Began adding refs and creating a bibtex file. Other light edits, too.
Still need to find out how to compile the bibtex file correctly.

@article{restifo2012adoptive,
  title={Adoptive immunotherapy for cancer: harnessing the T cell response},
  author={Restifo, Nicholas P and Dudley, Mark E and Rosenberg, Steven A},
  journal={Nature Reviews Immunology},
  volume={12},
  number={4},
  pages={269},
  year={2012},
  publisher={Nature Publishing Group}
}

@article{de2006paradoxical,
  title={Paradoxical roles of the immune system during cancer development},
  author={De Visser, Karin E and Eichten, Alexandra and Coussens, Lisa M},
  journal={Nature reviews cancer},
  volume={6},
  number={1},
  pages={24},
  year={2006},
  publisher={Nature Publishing Group}
}

@article{nieto2016emt,
  title={EMT: 2016},
  author={Nieto, M Angela and Huang, Ruby Yun-Ju and Jackson, Rebecca A and Thiery, Jean Paul},
  journal={Cell},
  volume={166},
  number={1},
  pages={21--45},
  year={2016},
  publisher={Elsevier}
}

@article{alderton2012metastasis,
  title={Metastasis: Epithelial to mesenchymal and back again},
  author={Alderton, Gemma K},
  journal={Nature Reviews Cancer},
  volume={13},
  number={1},
  pages={3},
  year={2012},
  publisher={Nature Publishing Group}
}

@article{terry2017new,
  title={New insights into the role of EMT in tumor immune escape},
  author={Terry, St{\'e}phane and Savagner, Pierre and Ortiz-Cuaran, Sandra and Mahjoubi, Linda and Saintigny, Pierre and Thiery, Jean-Paul and Chouaib, Salem},
  journal={Molecular oncology},
  volume={11},
  number={7},
  pages={824--846},
  year={2017},
  publisher={Wiley Online Library}
}

@article{hong2015ovol2,
  title={An Ovol2-Zeb1 mutual inhibitory circuit governs bidirectional and multi-step transition between epithelial and mesenchymal states},
  author={Hong, Tian and Watanabe, Kazuhide and Ta, Catherine Ha and Villarreal-Ponce, Alvaro and Nie, Qing and Dai, Xing},
  journal={PLoS computational biology},
  volume={11},
  number={11},
  pages={e1004569},
  year={2015},
  publisher={Public Library of Science}
}