### 2019-09-23 Paper/Supplement/Supplement_rewrite.tex
updated CleanPaper similarly and put LIHC in Supplement

\documentclass[11pt, a4paper, preprint]{article}
%\usepackage{outline}
%\usepackage{float}
%\usepackage{graphicx}
%\usepackage{color}
%\usepackage{amsmath}
%\usepackage{cite}
%\graphicspath{ {Figs/} }
%\usepackage{longtable}

\usepackage{amsmath, amssymb, graphicx, outline, float, cite, color}

%\input{../structure.tex} % Specifies the document structure and loads requires packages

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\newcommand{\tcr} { \textcolor{red} }

\title{Supplementary Materials for \\
Modeling the competing effects of the immune system and EMT on epithelial cancers
}
\author{Daniel R. Bergman$^{1}$,
Matthew Karikomi\,$^{1}$,
Qing Nie\,$^{1,2,*}$
and \\Adam L. MacLean\,$^{3,*}$
}
\affil{
  $^1$Department of Mathematics, University of California, Irvine,  Irvine, CA 92697, USA \\
  $^2$Department of Cell and Developmental Biology, University of California, Irvine, Irvine, CA 92697, USA \\
  $^3$Department of Biological Sciences, University of Southern California, Los Angeles, CA 90089, USA \\
  $^*$Correspondence:  qnie@uci.edu (Q.N.); macleana@usc.edu (A.L.M.)
}
\renewcommand\Authands{ and }

%\title{
%The competing effects of the immune system and EMT on epithelial cancers
%}

\date{}

\begin{document}
\maketitle
\tableofcontents

\section{Definition of parameters specifying the model}

\begin{table}[H]
\begin{center}
 \begin{tabular}{|| c | c||} 
 \hline
 Name & Description  \\ [0.5ex] 
 \hline
 $p$ & proliferation rate of tissue cells \\ 
 \hline
 $d_C$  & death rate of tissue cells \\
 \hline
$\Delta_\text{MIE}$ &  mesenchymal immune evasion \\
 \hline
 $\Delta_\text{MGA}$ & mesenchymal growth arrest    \\
 \hline
  $\Delta_A$ & mutant cells decreased apoptosis  \\
  \hline
  $\Delta_\text{IE}$ & mutant cells increased immune evasion  \\
  \hline
  $\Delta_P$ & mutant cells increased proliferation  \\
  \hline
 $K_0$ & EC50 term for negative feedback of tissue cells on own proliferation\\
 \hline
 $K_1$ & EC50 term for probability of NK cell finding mutant cell\\
 \hline
  $K_2$ & EC50 term for Treg inhibition of cytotoxic functions  \\
  \hline
  $K_3$ & EC50 term for how much TGF-$\beta$ each cell has \\
  \hline
  $K_4$ & EC50 term for TGF-$\beta$ activation of Tregs \\
  \hline
 $E_\text{NK}$ & rate of NKs clearing mutants  \\
  \hline
  $E_\text{CTL}$ & rate of CTLs clearing mutants \\
  \hline
  $\sigma_\text{NK}$ & NK source rate \\ 
  \hline
  $\sigma_\text{CTL}$ & CTL source rate per cleared mutant cell \\ 
  \hline
  $\sigma_\text{Treg}$ & Treg source rate per cleared mutant cell \\ 
  \hline
  $d_\text{NK}$ & NK death rate  \\ 
  \hline
  $d_\text{CTL}$ & CTL death rate \\ 
  \hline
  $d_\text{Treg}$ & Treg death rate \\ 
  \hline
  $k_\text{EMT}$ & EMT/MET rate  \\
  \hline
  $\sigma$ & standard deviation of noise in TGF-$\beta$ each cell receives  \\
  \hline
 $\tau_\text{max}$ & max amount of TGF-$\beta$ any cell can receive \\
  \hline 
 $\tau_\text{MUT}$ & rate of TGF-$\beta$ production by mutant cells\\
  \hline
 $\tau_\text{Treg}$ & rate of TGF-$\beta$ production by Treg\\
  \hline
\end{tabular}
  \caption{The model parameter names and descriptions. Note that many of these values are affected by the inflammation state of the system.}
\end{center}
\end{table}

\section{Parameter values used for simulation}

\begin{table}[H]
\begin{center}
 \begin{tabular}{|| c | p{7cm} | c | c ||} 
 \hline
 Name & Description & INFL Low Value & INFL High Value  \\ [0.5ex] 
 \hline\hline
  $p$ & weight of proliferation for tissue cells & 0.28 & \\ 
 \hline
 $d_C$  & weight of apoptosis for tissue cells & 0.14 & \\
 \hline
$\Delta_\text{MIE}$ & MIE & 0.6 & \\
 \hline
 $\Delta_\text{MGA}$ & MGA  & 0.2 &  \\
 \hline
  $\Delta_A$ & proportional decrease to weight of apoptosis for cells with mutated apoptosis pathway & 0.3 &  \\
  \hline
  $\Delta_\text{IE}$ & proportional increase to weight of immune evasion for cells with mutated immune evasion pathway & 0.48 &  \\
  \hline
  $\Delta_P$ & proportional increase to weight of proliferation for cells with mutated proliferation pathway & 0.36 &  \\
  \hline
 $K_0$ & EC50 term for negative feedback of tissue cells on own proliferation & 80 cells& \\
 \hline
 $K_1$ & EC50 term for probability of NK cell finding mutant cell & 8 cells& \\
 \hline
  $K_2$ & EC50 term for Treg inhibition of cytotoxic functions & 5 cells / volume & 0.025 cells / volume \\
  \hline
  $K_3$ & EC50 term for cumulative absorption of TGF-$\beta$ & 200 amount / volume & \\
  \hline
  $K_4$ & EC50 term for TGF-$\beta$ activation of Tregs & 50 amount / volume & \\
  \hline
 $E_\text{NK}$ & weight of NKs clearing mutants & 10 & 30  \\
  \hline
  $E_\text{CTL}$ & weight of CTLs clearing mutants & 200 & 600 \\
  \hline
  $\sigma_\text{NK}$ & NK source rate & 1.3 cells / cycle &  \\ 
  \hline
  $\sigma_\text{CTL}$ & CTL source rate per cleared mutant cell & 100 cells / (cleared mutants $\times$ cycles)& \\ 
  \hline
  $\sigma_\text{Treg}$ & Treg source rate per cleared mutant cell & 200 cells / (cleared mutants $\times$ concentration of TGF-$\beta$ $\times$ cycles)& \\ 
  \hline
  $d_\text{NK}$ & NK death rate & 0.13 / cycle &  \\ 
  \hline
  $d_\text{CTL}$ & CTL death rate & 0.0260 / cycle& \\ 
  \hline
  $d_\text{Treg}$ & Treg death rate & 0.0260 / cycle& \\ 
  \hline
  $k_\text{EMT}$ & EMT/MET rate & 0.01 / concentration of TGF-$\beta$ & \\
  \hline
  $\sigma$ & standard deviation of noise in TGF-$\beta$ each cell receives & 6 concentration of TGF-$\beta$ & \\
  \hline
 $\tau_\text{max}$ & max amount of TGF-$\beta$ any cell can receive & 500 concentration of TGF-$\beta$ &\\
  \hline 
 $\tau_\text{MUT}$ & rate of TGF-$\beta$ production by mutant cells & 0.05 concentration of TGF-$\beta$ / cell / cycle& \\
  \hline
 $\tau_\text{Treg}$ & rate of TGF-$\beta$ production by Treg & 0.5 concentration of TGF-$\beta$ / cell / cycle& \\
  \hline
  & RP Cancer Line  & 0.5 &  \\ 
 \hline
&  INFL High Duration & 30 cycles&   \\
 \hline
 & INFL Low Duration &30 cycles&   \\
 \hline
& Mes Threshold & 0.7 &    \\
 \hline
  & maximum initial mutation damage after warmup  & 0.01 &\\
  \hline
  & increase in probability to mutate for non-mutating proliferating cells & 0.0001 & \\
  \hline
\end{tabular}
  \caption{The model parameter names, descriptions, and values during both low and high inflammation. Parameters with only one value do not change with the inflammatory state.}
\end{center}
\end{table}

\newpage

\begin{table}[H]
\begin{center}
 \begin{tabular}{|| c | p{7cm} | c | c ||} 
 \hline
 Name & Description & INFL Low Value & INFL High Value  \\ [0.5ex] 
 \hline\hline
  $p$ & proliferation rate of tissue cells & 0.28 & \\ 
 \hline
 $d_C$  & death rate of tissue cells & 0.14 & \\
 \hline
$\Delta_\text{MIE}$ &  MIE & 0.6 & \\
 \hline
 $\Delta_\text{MGA}$ & MGA  & 0.2 &  \\
 \hline
  $\Delta_A$ & mutant cells decreased apoptosis & 0.3 &  \\
  \hline
  $\Delta_\text{IE}$ & mutant cells increased immune evasion & 0.48 &  \\
  \hline
  $\Delta_P$ & mutant cells increased proliferation & 0.36 &  \\
  \hline
 $K_0$ & EC50 term for negative feedback of tissue cells on own proliferation & 80 & \\
 \hline
 $K_1$ & EC50 term for probability of NK cell finding mutant cell & 8 & \\
 \hline
  $K_2$ & EC50 term for Treg inhibition of cytotoxic functions & 5 & 0.025  \\
  \hline
  $K_3$ & EC50 term for how much TGF-$\beta$ each cell has & 200 & \\
  \hline
  $K_4$ & EC50 term for TGF-$\beta$ activation of Tregs & 50 & \\
  \hline
 $E_\text{NK}$ & rate of NKs clearing mutants & 10 & 30  \\
  \hline
  $E_\text{CTL}$ & rate of CTLs clearing mutants & 200 & 600 \\
  \hline
  $\sigma_\text{NK}$ & NK source rate & 1.3 &  \\ 
  \hline
  $\sigma_\text{CTL}$ & CTL source rate per cleared mutant cell & 100 & \\ 
  \hline
  $\sigma_\text{Treg}$ & Treg source rate per cleared mutant cell & 200 & \\ 
  \hline
  $d_\text{NK}$ & NK death rate & 0.13 &  \\ 
  \hline
  $d_\text{CTL}$ & CTL death rate & 0.0260 & \\ 
  \hline
  $d_\text{Treg}$ & Treg death rate & 0.0260 & \\ 
  \hline
  $k_\text{EMT}$ & EMT/MET rate & 0.01 & \\
  \hline
  $\sigma$ & standard deviation of noise in TGF-$\beta$ each cell receives & 6 & \\
  \hline
 $\tau_\text{max}$ & max amount of TGF-$\beta$ any cell can receive & 500 &\\
  \hline 
 $\tau_\text{MUT}$ & rate of TGF-$\beta$ production by mutant cells & 0.05 & \\
  \hline
 $\tau_\text{Treg}$ & rate of TGF-$\beta$ production by Treg & 0.5 & \\
  \hline
  & RP Cancer Line  & 0.5 &  \\ 
 \hline
&  INFL High Duration & 30&   \\
 \hline
 & INFL Low Duration &30 &   \\
 \hline
& Mes Threshold & 0.7 &    \\
 \hline
  & maximum initial mutation damage after warmup  & 0.01 &\\
  \hline
  & increase in probability to mutate for non-mutating proliferating cells & 0.0001 & \\
  \hline
\end{tabular}
  \caption{The model parameter names, descriptions, and values during both low and high inflammation. Parameters with only one value do not change with the inflammatory state.}
\end{center}
\end{table}

\begin{figure}
\center
{\includegraphics[width=0.65\textwidth]{../Figs/Figure6/LIHC.pdf}}
\caption{A. K-means clustering of LIHC using gene ontology terms indicative of EMT and inflammation signatures ($k=2$).
B. Survival plots corresponding to the clustering on EMT and inflammation.
C. K-means clustering of LIHC using gene ontology terms indicative of an EMT signature ($k=2$).
D. Survival plots corresponding to the clustering on EMT.
E. K-means clustering of LIHC using gene ontology terms indicative of inflammation ($k=2$).
F. Survival plots corresponding to the clustering on inflammation.}
\label{fig:LIHC}
\end{figure}

%\newpage
%
%
% \begin{longtable}{||c | c | c | c | c | c | c | c | c | c ||} 
% \hline
% Name & Base Value & Single Patient & Survival Curve & Vary MIE & Vary MGA & Vary Treg TGFB & Vary MIE and INFL & Vary MGA and INFL & Heat Map\\ [0.5ex] 
% \hline\hline
% RP Cancer Line &  0.5 &  \\ 
% \hline
% MIE  & 0.6 & 0.9 & 0.4-0.9 & 0.7 & 0.5 & 0.9 \\
% \hline
% MGA & 0.2 & & 0.1-0.4 & 0 & 0.2 & 0 \\
% \hline
% INFL High Duration  & 30 & & & & 60 & & & & 60 \\
% \hline
%INFL Low Duration  & 30  \\
% \hline
%Mes Threshold & 0.7 \\
% \hline
% p  & 0.28 \\ 
% \hline
% d  & 0.14 \\
% \hline
% $N_{00}$  & 8\\
% \hline
% $N_0$  & 80 \\
% \hline
%  NK Efficacy Low  & 0.2 & 10 & & 10 & 10 & 10 & 10 & 10 & 10 \\
%  \hline
%  Treg EC50 Low & 5 & & & & & 2.5 \\
%  \hline
%  CTL Efficacy Low  & 4 & 200 & & 200 & 200 & 200 & 200 & 200 & 200\\
%  \hline
%  NK Efficacy Up & 2 & & & & 1.2 & & & &  1.2 \\
%  \hline
%  CTL Efficacy Up & 3 & & & & 3 & & & &  3\\
%  \hline
%  Treg Efficacy Up  & 200 & & & & 10& & & &  10\\
%  \hline
%  NK Source Low  & 1.3 \\ 
%  \hline
%  CTL Recruitment Low  & 100 \\ 
%  \hline
%  Treg Recruitment Low  & 200 \\
%  \hline
%  NK Recruitment Up & 1 & & & & & 3\\
%  \hline
%  CTL Recruitment Up & 1 & & & & & 3\\
%  \hline
%   Treg Recruitment Up & 1 & & & & & 5\\
%  \hline
%  $p_{mutation}$ Start & 1e-2 \\
%  \hline
%  $p_{mutation}$ & 1e-4 \\
%  \hline
%  Apoptosis Down & 0.3 \\
%  \hline
%  Immune Evasion & 0.48 \\
%  \hline
%  Proliferation Up & 0.36  \\
%  \hline
%  $k_{EMT}$  & .01 \\
%  \hline
%  c & 6 \\
%  \hline
%  TGFB Received EC50 & 200 \\
%  \hline
%  TGFB Max & 500 & 700 & & & & & &  & 700\\
%  \hline 
%  Mutant TGFB & 5e-2 & & & & & 0.5\\
%  \hline
%  Treg TGFB & 5e-1 \\
%  \hline
%\end{longtable}
%
%

\end{document}